EnVision Data Summary

### 2024-11-22 Christine Chan/EnVision CF/katie_dataset_check.qmd
basic AUC analysis for Katie

```{r setup}
#| include: false
library(tidyverse)
library(arsenal)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF")
```

```{r data cleaning}
# Import cleaned dataset
df <- read.csv("./Data_Clean/analysis_dataset.csv", na.strings = "")
# Set table options
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "range"),
    cat.stats = c("N", "Nmiss", "countpct"),
    digits = 1
  )
```

```{r results='asis'}
#| label: tbl-auc-numbers-overall
#| tbl-cap: Glucose and Insulin iAUC
tbl <- tableby(
  Diagnosis ~ iAUC30gluc + iAUC60gluc + iAUC120gluc + iAUC180gluc + iAUC30ins +
    iAUC60ins + iAUC120ins + iAUC180ins,
  data = df, control = mycontrols
)
summary(tbl, pfootnote = T)
```

```{r results='asis'}
#| label: fig-auc-hist
#| fig-cap: Glucose and Insulin iAUC Histograms
df %>%
  select(
    iAUC30gluc, iAUC60gluc, iAUC120gluc, iAUC180gluc, iAUC30ins, iAUC60ins,
    iAUC120ins, iAUC180ins
  ) %>%
  pivot_longer(everything()) %>%
  mutate(
    measure = factor(sub("iAUC\\d{2,3}", "", name),
      levels = c("gluc", "ins"), labels = c("Glucose", "Insulin")
    ),
    name = factor(sub("gluc|ins", "", name),
      levels = c("iAUC30", "iAUC60", "iAUC120", "iAUC180")
    )
  ) %>%
  ggplot(., aes(x = name, y = value)) +
  geom_boxplot() +
  facet_wrap(~measure)
```

>>>>

```{r setup}
#| include: false
library(tidyverse)
library(arsenal)
library(skimr)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF")
```

```{r data cleaning}
# Import cleaned dataset
df <- read.csv("./Data_Clean/envision_analysis_dataset.csv", na.strings = "")
# Find those missing both iAUC120ins and iAUC120gluc, and both iAUC180ins and
# iAUC180gluc
df$miss_120 <- factor(is.na(df$iAUC120gluc) & is.na(df$iAUC120ins),
  levels = c(F, T), labels = c("No", "Yes")
)
df$miss_180 <- factor(is.na(df$iAUC180gluc) & is.na(df$iAUC180ins),
  levels = c(F, T), labels = c("No", "Yes")
)
# Labels
mylabels <- list(
  miss_120 = "Missing iAUC120gluc and iAUC120ins?",
  miss_180 = "Missing iAUC180gluc and iAUC180ins?"
)
# Set table options
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "range"),
    cat.stats = c("N", "Nmiss", "countpct"),
    digits = 1
  )
```

# Demographics

```{r results='asis'}
#| label: tbl-demo
#| tbl-cap: Demographics by Participant
tbl <- tableby(
  ~ age_visit + height + weight + bmi + bmi_perc + origin_race + ethnicity +
    sex + CFTR,
  data = df %>% group_by(study_id) %>% slice_min(row_number()),
  control = mycontrols
)
summary(tbl, pfootnote = T)
```

```{r results='asis'}
#| label: tbl-demo-ogtt
#| tbl-cap: Demographics Treating OGTTs as Independent
tbl <- tableby(
  ~ age_visit + height + weight + bmi + bmi_perc + origin_race + ethnicity +
    sex + CFTR,
  data = df, control = mycontrols
)
summary(tbl, pfootnote = T)
```

# iAUC

```{r results='asis'}
#| label: tbl-auc-numbers-overall
#| tbl-cap: Glucose and Insulin iAUC
tbl <- tableby(
  ~ iAUC30gluc + iAUC60gluc + iAUC120gluc + iAUC180gluc + iAUC30ins +
    iAUC60ins + iAUC120ins + iAUC180ins + miss_120 + miss_180,
  data = df, control = mycontrols
)
summary(tbl, pfootnote = T, labelTranslations = mylabels)
```

```{r results='asis'}
#| label: tbl-auc-numbers-diagnosis
#| tbl-cap: Glucose and Insulin iAUC by Diagnosis
tbl <- tableby(
  Diagnosis ~ iAUC30gluc + iAUC60gluc + iAUC120gluc + iAUC180gluc + iAUC30ins +
    iAUC60ins + iAUC120ins + iAUC180ins + miss_120 + miss_180,
  data = df, control = mycontrols
)
summary(tbl, pfootnote = T, labelTranslations = mylabels)
```

```{r results='asis'}
#| label: fig-auc-hist
#| fig-cap: Glucose and Insulin iAUC Boxplots
df %>%
  select(
    iAUC30gluc, iAUC60gluc, iAUC120gluc, iAUC180gluc, iAUC30ins, iAUC60ins,
    iAUC120ins, iAUC180ins
  ) %>%
  pivot_longer(everything()) %>%
  mutate(
    measure = factor(sub("iAUC\\d{2,3}", "", name),
      levels = c("gluc", "ins"), labels = c("Glucose", "Insulin")
    ),
    name = factor(sub("gluc|ins", "", name),
      levels = c("iAUC30", "iAUC60", "iAUC120", "iAUC180")
    )
  ) %>%
  filter(!is.na(value)) %>%
  ggplot(., aes(x = name, y = value)) +
  geom_boxplot() +
  facet_wrap(~measure)
```

# EDA

```{r}
#| label: fig-skimr
#| fig-cap: Data Skim
skim(df)
```
